---
title: "Untitled"
author: "Dr. Jose Mawyin"
date: "12/10/2020"
output:
  prettydoc::html_pretty:
    theme: architec
    toc: true
    #toc_float: true
    fig_caption: true
    toc_collapsed: false
    toc_depth: 5
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache=TRUE)
#install.packages("beepr")
#install.packages(c("VIM", "rpart", "randomForest", "party", "gbm", "Cubist", "rpart.plot"))
#install.packages("PerformanceAnalytics")

library(tidyr)

library(VIM)
library(Hmisc)
#Corr Plot
library(corrplot)
library(RColorBrewer)
library(ggcorrplot)
library(PerformanceAnalytics)
library(magrittr)
library(reshape2)
library(ggplot2)

#Data Imputation
library(mice)

#Train model
library(caret)

#Timing
library(tictoc)

#Tree Based Plots
library(rpart)
library(randomForest)
library(party)

library(gbm)
#library(pryr)
library(Cubist)
library(Hmisc)
library(mice)
library(magrittr)
#library(partykit)
library(rpart.plot)

library(beepr)

#Parallel Processing
library(doParallel)
cl <- makePSOCKcluster(4)
registerDoParallel(cl)
#stopCluster(cl);registerDoSEQ()
set.seed(1978)
```

```{r, eval = TRUE}
hist.data.frame <- function(x, ..., colors=rainbow(ncol(x))) {
    col<-1
    hist<-function(...) {
        graphics::hist(..., col=colors[col])
        col <<- col+1
    }
    f <- Hmisc:::hist.data.frame
    environment(f) <- environment()
    f(x,...)
}
```

## Predicting District Yield


The following data frames contain per region weather parameters at a half hour resolution for a 7 day period and for all 300+ districts.
```{r Data_read}
data_loc <- '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/Predictor_Response_11-11to11-18.csv'
data_loc2 <- '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/Predictor_Response_11-18to11-25.csv'

data_loc3 <- '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/Predictor_Response_11-25to11-30.csv'
pv_ac_weather_df <- read.csv(file = data_loc)
pv_ac_weather_df_2 <- read.csv(file = data_loc2)
pv_ac_weather_df_3 <- read.csv(file = data_loc3)
dim(pv_ac_weather_df); dim(pv_ac_weather_df_2); dim(pv_ac_weather_df_3)
```



```{r pressure, echo=FALSE}
head(pv_ac_weather_df)
dim(pv_ac_weather_df)
#pv_ac_weather_df_s <- sample(pv_ac_weather_df, 1000)
pv_ac_weather_df_s <- pv_ac_weather_df[sample(nrow(pv_ac_weather_df), 10000, replace = FALSE, prob = NULL),]
dim(pv_ac_weather_df_s)
```

```{r}
pv_ac_weather_columns <- c("hourly_AC_Power","temp_air", "wind_speed", "total_clouds", "low_clouds","mid_clouds", "high_clouds","Latitude","Longitude","Daily_Hour","Day_of_Year")
#pv_ac_weather_columns <- c("hourly_AC_Power","temp_air", "wind_speed", "ghi", "dni","dhi", "total_clouds", "low_clouds","mid_clouds", "high_clouds","Latitude","Longitude","poa_global","poa_direct","poa_diffuse","poa_sky_diffuse","poa_ground_diffuse","Cell_temp","Daily_Hour","Day_of_Year")
pv_ac_weather_reduced <- pv_ac_weather_df_s[pv_ac_weather_columns]
head(pv_ac_weather_reduced)
```

```{r}
colSums(is.na(pv_ac_weather_reduced))
```
```{r, eval = TRUE}
hist(pv_ac_weather_reduced[,2:10])
```

```{r}
corr <- model.matrix(~0+., data=pv_ac_weather_reduced) %>% 
  cor(use="pairwise.complete.obs")


ggplot(melt(corr), aes(Var1, Var2, fill=value)) +
  geom_tile(height=0.8, width=0.8) +
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  coord_equal() +
  labs(x="",y="",fill="Corr") +
  theme(axis.text.x=element_text(size=8, angle=90, vjust=1, hjust=1, 
                                 margin=ggplot2::margin(-3,0,0,0)),
        axis.text.y=element_text(size=8, margin=ggplot2::margin(0,-3,0,0)),
        panel.grid.major=element_blank()) +
  ggtitle("Correlation Plot of Variables in District Yield Prediction") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
df <- pv_ac_weather_reduced


#Creating Test and Training Set
trainIndex <- createDataPartition(df$hourly_AC_Power, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_Train <- df[ trainIndex,]
df_Test  <- df[-trainIndex,]
```


```{r}
tic()
#Linear Regression
lm_model <- train(
  hourly_AC_Power~., data = df_Train, method = "lm")

cat("\nLinear Regression training time:\n")
toc()
lm_predictions <- lm_model %>% predict(df_Test)
# Model performance metrics
lm_accuracy <- data.frame( 
  Model = "Linear Regression",
  RMSE = caret::RMSE(lm_predictions, df_Test$hourly_AC_Power),
  Rsquare = caret::R2(lm_predictions, df_Test$hourly_AC_Power)
)
beep(2)
#==================
#knn
tic()
knnModel <- train(hourly_AC_Power~., data = df_Train,
                  method = "knn",
                  preProc = c("center", "scale"),
                  tuneLength = 10)
cat("\nknn training time:\n")
toc();beep(2)
knnModel_predictions <- knnModel %>% predict(df_Test)
# Model performance metrics
knn_Accuracy <- data.frame(
  Model = "k-Nearest Neighbors",
  RMSE = caret::RMSE(knnModel_predictions,df_Test$hourly_AC_Power),
  Rsquare = caret::R2(knnModel_predictions,df_Test$hourly_AC_Power))
#==================
tic()
metric <- "RMSE"
control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")

tunegrid <- expand.grid(.mtry=c(1:15))
rf_grid <- train(hourly_AC_Power ~ ., data = df_Train, 
                       method="rf", metric=metric, tuneGrid=tunegrid, trControl=control)
beep('ping')
#Random Forest predictions
rf_predictions <- rf_grid %>% predict(df_Test)
# Model performance metrics
rf_accuracy <- data.frame( 
  Model = "Random Forest Grid",
  RMSE = caret::RMSE(rf_predictions, df_Test$hourly_AC_Power),
  Rsquare = caret::R2(rf_predictions, df_Test$hourly_AC_Power)
)
cat("\nRandom Forest training time:\n")
toc()
beep(2)
```




```{r}
all_model_accuracy <- rbind(lm_accuracy,knn_Accuracy,rf_accuracy)
all_model_accuracy
```

```{r}
All_predictions <- cbind(lm_predictions,knnModel_predictions,rf_predictions) %>% as.data.frame()
All_predictions <- cbind(All_predictions,df_Test$hourly_AC_Power)
names(All_predictions)[names(All_predictions) == 'df_Test$hourly_AC_Power'] <- "Hourly_AC_Power"
head(All_predictions)
#Melting all column values together using PH as index
df.m <- melt(All_predictions, "Hourly_AC_Power")
names(df.m)[names(df.m) == 'variable'] <- "Model"
head(df.m)

ggplot(df.m, aes(value, Hourly_AC_Power,colour = Model)) + 
  geom_point() + 
  facet_wrap(~Model, scales = "free", ncol = 2) +
  geom_abline() +coord_cartesian(xlim = c(-0.1, 0.4), ylim = c(-0.1, 0.4)) + #stat_density_2d(aes(fill = ..level..), geom="polygon")+
  ggtitle("Predicted Vs Observed Solar Yield (Watt/Watt-peak) - 10000 Samples")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(df.m, aes(value, Hourly_AC_Power,colour = Model)) + 
  geom_point() + 
  facet_wrap(~Model, scales = "free", ncol = 2) +
  geom_abline() +coord_cartesian(xlim = c(-0.1, 0.4), ylim = c(-0.1, 0.4)) + #stat_density_2d(aes(fill = ..level..), geom="polygon")+
  ggtitle("Predicted Vs Observed Solar Yield (Watt/Watt-peak) - 1000 Samples")+
  theme(plot.title = element_text(hjust = 0.5))
```


## Predicting Regional Generation Output


Name of Regions in the District Yield Data Set:

"North East"               "North West"               "Yorkshire and The Humber" "East Midlands"           
 "West Midlands"            "South West"               "East of England"          "South East"              "London"   
 
 Name of Regions in the Region Generation Output Data Set:
 
 "NATIONAL"              "Eastern England"       "East Midlands"         "London"                "West Midlands"        
 "North Eastern England" "North Western England" "South Eastern England" "Southern England"      "South Western England"
"Yorkshire"  


```{r}
Region_name <- "North West"    
#============= Adding District Yields
pv_ac_region_reduced <- pv_ac_weather_df
pv_ac_region_reduced <- filter(pv_ac_region_reduced, Region == Region_name)
pv_ac_region_columns <- c("TIMESTAMP","District",  "hourly_AC_Power")
pv_ac_region_reduced <- pv_ac_region_reduced[pv_ac_region_columns]
#head(pv_ac_region_reduced)
pv_ac_region_reduced <- spread(pv_ac_region_reduced, District, hourly_AC_Power)
one_week_pv_ac_region_reduced <- pv_ac_region_reduced
#head(pv_ac_region_reduced); dim(pv_ac_region_reduced)
#===========
pv_ac_region_reduced_2 <- pv_ac_weather_df_2
pv_ac_region_reduced_2 <- filter(pv_ac_region_reduced_2, Region == Region_name)
pv_ac_region_columns <- c("TIMESTAMP","District",  "hourly_AC_Power")
pv_ac_region_reduced_2 <- pv_ac_region_reduced_2[pv_ac_region_columns]
#head(pv_ac_region_reduced)
pv_ac_region_reduced_2 <- spread(pv_ac_region_reduced_2, District, hourly_AC_Power)
#head(pv_ac_region_reduced_2); dim(pv_ac_region_reduced_2)
two_weeks_pv_ac_region_reduced <- rbind(pv_ac_region_reduced,pv_ac_region_reduced_2)
#===========
pv_ac_region_reduced_3 <- pv_ac_weather_df_3
pv_ac_region_reduced_3 <- filter(pv_ac_region_reduced_3, Region == Region_name)
pv_ac_region_columns <- c("TIMESTAMP","District",  "hourly_AC_Power")
pv_ac_region_reduced_3 <- pv_ac_region_reduced_3[pv_ac_region_columns]
#head(pv_ac_region_reduced)
pv_ac_region_reduced_3 <- spread(pv_ac_region_reduced_3, District, hourly_AC_Power)
#============= Adding Regional Yields
Region_name_Obs <- "North Western England"
Shef_SS_region_output_loc = '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/All_region_daily_ob_11_11_to_11_18.csv'
Shef_SS_region_output <- read.csv(file = Shef_SS_region_output_loc)
Shef_SS_region_output_loc2 = '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/All_region_daily_ob_11_18_to_11_25.csv'
Shef_SS_region_output_2 <- read.csv(file = Shef_SS_region_output_loc2)
Shef_SS_region_output_loc3 = '/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /698/Data/All_region_daily_ob_11_25_to_11_30.csv'
Shef_SS_region_output_3 <- read.csv(file = Shef_SS_region_output_loc3)
#Shef_SS_region_output <- rbind(Shef_SS_region_output,Shef_SS_region_output_2)
Shef_SS_region_output <- filter(Shef_SS_region_output, Region_name == Region_name_Obs)
one_week_Shef_SS_region_output <- Shef_SS_region_output
#=======
Shef_SS_region_output_2 <- filter(Shef_SS_region_output_2, Region_name == Region_name_Obs)
two_weeks_Shef_SS_region_output <- rbind(Shef_SS_region_output,Shef_SS_region_output_2)
#=======
Shef_SS_region_output_3 <- filter(Shef_SS_region_output_3, Region_name == Region_name_Obs)
dim(one_week_pv_ac_region_reduced); dim(one_week_Shef_SS_region_output)

#=============
train_df <- one_week_pv_ac_region_reduced[,-1] ;train_df$Region_gen <- one_week_Shef_SS_region_output$Generation
#train_df <- two_weeks_pv_ac_region_reduced[,-1] ;train_df$Region_gen <- two_weeks_Shef_SS_region_output$Generation
#head(train_df)
df <- train_df


#Creating Test and Training Set
trainIndex <- createDataPartition(df$Region_gen, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_Train <- df[ trainIndex,]
df_Test  <- df[-trainIndex,]

#========
tic()
#Linear Regression
lm_model <- train(
  Region_gen~., data = df_Train, method = "lm")

cat("\nLinear Regression training time:\n")
toc()
lm_predictions <- lm_model %>% predict(df_Test)
# Model performance metrics
lm_accuracy <- data.frame( 
  Model = "Linear Regression",
  RMSE = caret::RMSE(lm_predictions, df_Test$Region_gen),
  Rsquare = caret::R2(lm_predictions, df_Test$Region_gen)
)
#beep(2)
#==================
#knn
tic()
knnModel <- train(Region_gen~., data = df_Train,
                  method = "knn",
                  preProc = c("center", "scale"),
                  tuneLength = 10)
cat("\nknn training time:\n")
toc();#beep(2)
knnModel_predictions <- knnModel %>% predict(df_Test)
# Model performance metrics
knn_Accuracy <- data.frame(
  Model = "k-Nearest Neighbors",
  RMSE = caret::RMSE(knnModel_predictions,df_Test$Region_gen),
  Rsquare = caret::R2(knnModel_predictions,df_Test$Region_gen))
#==================
tic()
metric <- "RMSE"
control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")

tunegrid <- expand.grid(.mtry=c(1:15))
rf_grid <- train(Region_gen ~ ., data = df_Train, 
                       method="rf", metric=metric, tuneGrid=tunegrid, trControl=control)
#beep('ping')
#Random Forest predictions
rf_predictions <- rf_grid %>% predict(df_Test)
# Model performance metrics
rf_accuracy <- data.frame( 
  Model = "Random Forest Grid",
  RMSE = caret::RMSE(rf_predictions, df_Test$Region_gen),
  Rsquare = caret::R2(rf_predictions, df_Test$Region_gen)
)
cat("\nRandom Forest training time:\n")
toc()
#beep(2)

all_model_accuracy <- rbind(lm_accuracy,knn_Accuracy,rf_accuracy)
all_model_accuracy

All_predictions <- cbind(lm_predictions,knnModel_predictions,rf_predictions) %>% as.data.frame()
All_predictions <- cbind(All_predictions,df_Test$Region_gen)
names(All_predictions)[names(All_predictions) == 'df_Test$Region_gen'] <- "Region_gen"
#head(All_predictions)
#Melting all column values together using PH as index
df.m <- melt(All_predictions, "Region_gen")
names(df.m)[names(df.m) == 'variable'] <- "Model"
#head(df.m)

ggplot(df.m, aes(value, Region_gen,colour = Model)) + 
  geom_point() + 
  facet_wrap(~Model, scales = "free", ncol = 2) +
  geom_abline() +#coord_cartesian(xlim = c(0, 700), ylim = c(0, 700)) + #stat_density_2d(aes(fill = ..level..), geom="polygon")+
  ggtitle("Predicted Vs Observed Solar Yield (Watt/Watt-peak) - 10000 Samples")+
  theme(plot.title = element_text(hjust = 0.5))

rf_predictions_next_week <- rf_grid %>% predict(pv_ac_region_reduced_2)
plot(rf_predictions_next_week, main = paste(Region_name , "Region : Observed (Green) Vs. Predicted (Red)"), ylab = 'Generation (GWh)',xlab = 'Time',col="red")
lines(Shef_SS_region_output_2$Generation,col="green")
beep(4)
```


Linear Regression	3.715098e+07	0.001017149		
k-Nearest Neighbors	1.211324e+01	0.881977543		
Random Forest Grid	1.481217e+01	0.861682798

Linear Regression	5906.407254	0.003484721		
k-Nearest Neighbors	8.456962	0.945825033		
Random Forest Grid	8.327947	0.936611559	



